---
title: "Berlin Tempelhof Temperature Forecasting with ARIMA"
output: 
    md_document:
        variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r ImportLibraries, message=F, warning=F}
library(lubridate)
library(dplyr)
library(ggplot2)
library(forecast)
library(tseries)
```

### Import and Cleanup Data


```{r ImportData}
tempelhof=read.table("TG_STAID002759.txt", skip=20, 
                     col.names=c("SOUID", "date", "temp", "Q_TG"), sep=',')

tempelhof= tempelhof %>% select(date, temp) %>% mutate(date=ymd(date)) %>% mutate(temp=temp/10) %>% 
    filter(between(date, as.Date("2010-01-01"), as.Date("2020-04-30")))

head(tempelhof, 10)
```

**check data looks similiar as in python**

```{r TempGraph}
tempelhof2=ts(tempelhof$temp, start=c(2010,1), frequency=365)

autoplot(tempelhof2)+ ggtitle("Daily Berlin Tempelhof Temperature 2010-2020") + 
    theme(plot.title = element_text(hjust = 0.5))
```

### Stationarize Data

* removes seasonality (`seassadj()`) and trend ('diff()') from decomposed (`stl()`) data

```{r stationarize}
stationed=tempelhof2 %>% stl(s.window='periodic') %>% seasadj() %>% diff()

ggtsdisplay(stationed, main="", lag.max=60)
```

### Grid Search for Hyperparameter Optimization

```{r gridsearch}
ARIMAfit=auto.arima(stationed, approximation=F, trace=F)

summary(ARIMAfit)
```

```{r modelfit}
model=Arima(stationed, order=c(5,0,0))

checkresiduals(model)
```

**Based on the [documentation](https://otexts.com/fpp2/arima-r.html) the main take away point in the residual plot is that the residuals are not all white noise since not all autocorrelations are within threshold limits in ACF plot.**

### Forecasting 90 days into future 

* slight improvement to python 
* below show the forecast is ineffective after the 5th day

```{r modelfit2, message=F, warning=F}
forecast=data.frame(forecast(model, h=90))

head(forecast$Point.Forecast, 15)

autoplot(forecast(model, h=90)) + xlim(c(2020.0, 2020.4))
```



